---
title: "Reproducible documents"
author: "Alicia Muñoz Espinosa"
format: html
editor: visual
---

RANDOM SENTENCE SOMEWHERE NEAR THE TOP OF THE FILE (BELOW THE YAML HEADER)

ANOTHER RANDOM SENTENCE SOMEWHERE NEAR THE TOP OF THE FILE

```{r setup}
library(tidyverse)
```

## "About me"

- Alicia

- UAEIS

**Researcher** with focus in *Public Health*

## "Simple code"

```{r}
3 * 3
```

```{r}
2 + 2
```

```{r}
#| message: false
#| warning: false
library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |> 
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |> 
  filter(glykemi == 0) |> 
  select(-glykemi)

tidied_post_meal_Data <- tidier_post_meal_data |> 
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )

```

## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI))
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

 

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

## Excercise: discrete plots

```{r}
#| fig-cap: "Personas con medición de glicemia"
#| label: fig-conteo-glicemia
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r}
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap: 
#| - "Distribution of BMI."
#| - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap: 
  #| - "Distribution of BMI."
  #| - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose."
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)
```

## Select helpers

```{r}
select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming columns

```{r}
small_post_meal_data |> 
  rename(
    auc_glucose = auc_pg, 
    auc_insulin = auc_ins
    )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |> 
  select(contains("OGTT"))

# Task 2. 
post_meal_data |> 
  select(contains(".."), - contains("..."))

# Task 3. 
post_meal_data |> 
  rename(
    P.Glucose_5.OGTT = P.Glucose..5.OGTT,
    Insulin_5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
post_meal_data |> 
  select(id, BMI, Length, Age) |> 
  rename(
    Height = Length
  )
```

## Filtering data

```{r}
small_post_meal_data |> 
  filter(glykemi == 1)

small_post_meal_data |> 
  filter(BMI >= 25)

small_post_meal_data |> 
  filter(Group == "CTR")

small_post_meal_data |> 
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |> 
  filter(Group == "CTR" | BMI >= 25)
```

## Modifying columns

```{r}
tidier_post_meal_data |> 
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
    )
```

## Practice with filter and mutate

```{r}
# Task 1.
post_meal_data |> 
  filter(BMI >= 25 | BMI <=40 & Group == "FDR")

# Task 2.
post_meal_data |> 
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |> 
  filter(percent_body_fat > 30)
```

## Visualizing with a jitter plot

```{r}
tidied_post_meal_Data |> 
  mutate(overweight = if_else(BMI >= 25, "Overweight", "Not overweight")) |> 
  ggplot(aes(x = overweight, y = insulin_glucose_ratio)) +
  geom_jitter()
```

```{r}
tidied_post_meal_Data |> 
  filter(BMI < 30) |> 
  ggplot(aes(x = Age, y = auc_insulin, colour = Group)) +
  geom_point()
```

```{r}
tidied_post_meal_Data |> 
  filter(BMI < 30) |> 
  mutate(young = if_else(Age < 40, "Young", "Old")) |> 
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(cols = vars(young))
```

```{r}
tidied_post_meal_Data |> 
  filter(BMI < 30) |> 
  mutate(young = if_else(Age < 40, "Young", "Old")) |> 
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young))
```

```{r}
tidied_post_meal_Data |> 
  filter(BMI < 30) |> 
  mutate(young = if_else(Age < 40, "Young", "Old")) |> 
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(rows = vars(young)) +
  labs(
    x = "Body Mass Index (BMI)",
    y = "Insulin to glucose ratio"
  )
```

## Make multidimensional plots!

```{r}
#| label: fig-multidimensional
#| fig-cap: "Two, multidimensional plots of our data"
#| fig-subcap: 
#|    - "Distribution of percent body fat by group and age category"
#|    - "Relationship between BMI and bone mineral density, split by insulin sensitivity"
#| layout-ncol: 2
# Task 1.
post_meal_data |> 
  mutate(
    percent_body_fat = (Fat.mass.DXA/Weight) * 100,
    young = if_else(Age < 40, "Young", "Old")
    ) |> 
ggplot(aes(x = Group, y = percent_body_fat, colour = young)) +
geom_jitter(width = 0.2)

#Task 2.
post_meal_data |> 
  mutate(
    insulin_glucose_ratio = (auc_ins / auc_pg),
    insulin_insensitive = if_else(insulin_glucose_ratio > mean(insulin_glucose_ratio), "Insensitive", "Sensitive" )
  ) |> 
ggplot(aes(x = BMI, y = Bone.mineral.DXA, colour = insulin_insensitive)) +
geom_point() +
geom_smooth() +
facet_grid(rows = vars(insulin_insensitive))
```

## Changing colours

```{r}
#Tasks 1. and 2.
post_meal_data |> 
  mutate(hyperglycemia = if_else(glykemi == 1, "Yes", "No")) |> 
  ggplot(aes(x = Group, fill = hyperglycemia)) +
  geom_bar(position = position_dodge()) +
  scale_fill_viridis_d(option = "cividis")

#Tasks 3. and 4.
post_meal_data |> 
  ggplot(aes(x = auc_pg, y = auc_ins, colour = BMI)) +
  geom_point() +
  scale_colour_viridis_c(option = "inferno")
```

## Calculating summary statistics

```{r}
tidied_post_meal_Data |> 
  summarise(max_bmi = max(BMI, na.rm = TRUE),
            min_bmi = min(BMI, na.rm = TRUE)
  )
```

## Summarise with group by

```{r}
tidied_post_meal_Data|> 
  group_by(Group) |> 
  summarise(
    mean_age = mean(Age),
    mean_bmi = mean(BMI)
  ) |> 
  ungroup() |> 
  knitr::kable()
```

```{r}
tidied_post_meal_Data|> 
  group_by(Group) |> 
  summarise(
    mean_age = round(mean(Age), 1),
    mean_bmi = round(mean(BMI), 1)
  ) |> 
  ungroup() |> 
  knitr::kable()
```

```{r}
tidied_post_meal_Data|> 
  group_by(Group) |> 
  summarise(
    "Mean Age (years)" = round(mean(Age), 1),
    "Mean BMI (Kg/m^2^)" = round(mean(BMI), 1)
  ) |> 
  ungroup() |> 
  knitr::kable()
```

```{r}
#| label: tbl-mean-age-bmi
#| tbl-cap: "Mean values of Age and BMI for each group."
tidied_post_meal_Data|> 
  group_by(Group) |> 
  summarise(
    "Mean Age (years)" = round(mean(Age), 1),
    "Mean BMI (Kg/m^2^)" = round(mean(BMI), 1)
  ) |> 
  ungroup() |> 
  knitr::kable()
```
