---
title: "Reproducible documents"
author: "Alicia Muñoz Espinosa"
format: html
editor: visual
---

RANDOM SENTENCE SOMEWHERE NEAR THE TOP OF THE FILE (BELOW THE YAML HEADER)

ANOTHER RANDOM SENTENCE SOMEWHERE NEAR THE TOP OF THE FILE

```{r setup}
library(tidyverse)
```

## "About me"

- Alicia

- UAEIS

**Researcher** with focus in *Public Health*

## "Simple code"

```{r}
3 * 3
```

```{r}
2 + 2
```

```{r}
#| message: false
#| warning: false
library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |> 
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |> 
  filter(glykemi == 0) |> 
  select(-glykemi)

tidied_post_meal_Data <- tidier_post_meal_data |> 
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )

```

## Showing the data

```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI))
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

 

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

## Excercise: discrete plots

```{r}
#| fig-cap: "Personas con medición de glicemia"
#| label: fig-conteo-glicemia
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r}
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap: 
#| - "Distribution of BMI."
#| - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap: 
  #| - "Distribution of BMI."
  #| - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continous variables

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose."
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data

```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)
```

## Select helpers

```{r}
select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```

## Renaming columns 

```{r}
small_post_meal_data |> 
  rename(
    auc_glucose = auc_pg, 
    auc_insulin = auc_ins
    )
```

## Select and rename tasks

```{r}
# Task 1.
post_meal_data |> 
  select(contains("OGTT"))

# Task 2. 
post_meal_data |> 
  select(contains(".."), - contains("..."))

# Task 3. 
post_meal_data |> 
  rename(
    P.Glucose_5.OGTT = P.Glucose..5.OGTT,
    Insulin_5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
post_meal_data |> 
  select(id, BMI, Length, Age) |> 
  rename(
    Height = Length
  )
```

## Filtering data

```{r}
small_post_meal_data |> 
  filter(glykemi == 1)

small_post_meal_data |> 
  filter(BMI >= 25)

small_post_meal_data |> 
  filter(Group == "CTR")

small_post_meal_data |> 
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |> 
  filter(Group == "CTR" | BMI >= 25)
```

## Modifying columns

```{r}
tidier_post_meal_data |> 
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
    )
```
